### 2022-05-29 final-data-analysis.Rmd
tidying

---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
house <- read.csv("/Users/atticusw/Desktop/pton-market-data.csv")

# Tidying up
house <- house %>% 
  mutate(Price = strtoi(str_replace_all(str_sub(Sold.Price, 2, -4), ",", ""))) %>%
  rename(nbhd = Neighborhood, bed = Bed.Rooms, fullBath = Full.Baths, halfBath = Half.Baths, style = Style, age = Year.Built) %>%
  rename(lotSize = Lot.Size, lastPrice = Last.Price, originalPrice = Original.Price, soldPrice = Price) %>%
  select(- Sold.Price) %>%
  rename(marketDays = Days.on.Market) %>%
  separate(col = Sold.Date, into = c("monthSold", "daySold", "yearSold"), sep = "/") %>%
  mutate(yearSold = strtoi(yearSold) + 2000) %>%
  mutate(daySold = strtoi(daySold), monthSold = strtoi(monthSold)) %>%
  mutate(age = yearSold - strtoi(age)) %>%
  select(- Property.Marketing.Period) %>%
  mutate(originalPrice = strtoi(str_replace_all(str_sub(originalPrice, 2, -4), ",", ""))) %>%
  mutate(lastPrice = strtoi(str_replace_all(str_sub(lastPrice, 2, -4), ",", ""))) %>%
  select(- X) %>%
  mutate(lotSize = as.double(lotSize)) %>%
  select(- Address) %>%
  mutate(marketDays = strtoi(marketDays))
```

Response variable: soldPrice, (originalPrice)

Predictors: nbhd (neighborhood), bed, fullBath, halfBath, style, age, (originalPrice), (lastPrice), marketDays, yearSold, daySold, monthSold

Regression methods we will use: 
- full linear model (lm)
- subset selection (sub)
- ridge (ridge)
- lasso (lasso)
- principal components (pcr)

```{r}
# some final tidying

house <- house %>%
  select(- lotSize)  #too many NAs

house <- house %>%
  mutate(nbhd = replace(nbhd, nbhd == "Battelfield Area", "Battlefield Area"), nbhd = replace(nbhd, nbhd == "princeton Ridge", "Princeton Ridge"), nbhd = replace(nbhd, nbhd == "LIttlebrook", "Littlebrook"), nbhd = replace(nbhd, nbhd == "griggs Farm", "Griggs Farm"), nbhd = replace(nbhd, nbhd == "PrettyBrook Area", "Pretty Brook Area"), nbhd = replace(nbhd, nbhd == "Palmer Sq", "Palmer Square"), nbhd = replace(nbhd, nbhd == "Institute", "Institute Area"), nbhd = replace(nbhd, nbhd == "Rriverside", "Riverside"), nbhd = replace(nbhd, nbhd == "Rriverside", "Riverside"))  #fix area names

house <- house %>%
  filter(! nbhd %in% c("Edgerstoune", "Northridge", "Rushbrook", "Russell Estates", "The Preserve", "Governors Lane", "Princeton Borough", "Preserve", "Queenston")) # delete areas with at most 3 houses

house <- house %>%
  filter(! style %in% badStyles$style) # delete styles with at most 3 houses
```

```{r}

```