### 2022-05-30 paper-writeup.Rmd
paper-writeup initialize

---
title: "Data Science Final Project"
output:
  pdf_document: default
  html_notebook: default
author: Brandon Feder, Atticus Wang
date: May 30, 2022
---