### 2022-05-30 paper-writeup.Rmd

---
title: "Data Science Final Project"
output:
  pdf_document: default
  html_notebook: default
author: Brandon Feder, Atticus Wang
date: May 30, 2022
---

# Introduction

---OUR PLAN---

Response variable: soldPrice (price at which the house was sold)

Possible predictors:

- nbhd (neighborhood) 
- bed (number of bedrooms) 
- fullBath (number of full baths) 
- halfBath (number of half baths) 
- style (style) 
- age (yearSold minus yearBuilt) 
- originalPrice (original price) 
- lastPrice (price at which the house was last sold) 
- marketDays (days the house was on market) 
- yearSold, daySold, monthSold (date at which the house was sold)

Regression methods we will use:

- full linear model (lm) 
- subset selection (forward, backward) 
- ridge (ridge) 
- lasso (lasso) 
- elastic net (ela) 
- principal components regression (pcr)

# Regression Analysis

Our final models are Lasso and PCR. Lasso has the lowest variance, while PCR has the highest R2.

# Discussion and Limitations

Assumptions possibly not met, curse of dimensionality, did not incorporate census data

# Conclusion

# Additional Work

Introduce the other models we tried