### 2022-05-30 paper-writeup.Rmd
update paper

---
title: "Data Science Final Project"
output:
  pdf_document: default
  html_notebook: default
author: Brandon Feder, Atticus Wang
date: May 30, 2022
---

# Introduction

## Research question and hypothesis

Princeton has many diverse houses ranging in size, location, age, and many other factors. With this comes a huge range of house prices. But how do the characteristics of a house determine its price? Our research studies how factors such as the number of bedrooms, location, and the year built affect the price of the house. More precisely, we try to predict `soldPrice`, the price at which a house is sold, using the following possible predictors:

- nbhd (neighborhood) 
- bed (number of bedrooms) 
- fullBath (number of full baths) 
- halfBath (number of half baths) 
- style (style) 
- age (yearSold minus yearBuilt) 
- marketDays (days the house was on market) 
- yearSold, daySold, monthSold (date at which the house was sold)

Based off our own experience, we predict that neighborhood/address and the year built have a significant impact on the price of the house.

## Data description

Beatrice Bloom, a Princeton Residential Specialist, provides many great resources about the Princeton housing market including a table of houses sold in Princeton since 2011. This data can be found [here](https://www.realestate-princeton.com/market-analysis/princeton-pending/). We intend to use this data to answer our question. The data is stored in `./data/pton-market-data.csv` in the Github repo.

## Exploratory data analysis

Use `group_by` and `summarize` to find the top-10 styles and neighborhoods with the highest price:

```{r, echo = FALSE, warning = FALSE, message = FALSE}
library(tidyverse)
house <- read.csv("./data/pton-market-data.csv")

#Tidying up

house <- house %>% 
  mutate(Price = strtoi(str_replace_all(str_sub(Sold.Price, 2, -4), ",", ""))) %>%
  rename(nbhd = Neighborhood, 
         bed = Bed.Rooms, 
         fullBath = Full.Baths, 
         halfBath = Half.Baths, 
         style = Style, 
         age = Year.Built) %>%
  rename(lotSize = Lot.Size, 
         lastPrice = Last.Price, 
         originalPrice = Original.Price, 
         soldPrice = Price) %>%
  select(- Sold.Price) %>%
  rename(marketDays = Days.on.Market) %>%
  separate(col = Sold.Date, into = c("monthSold", "daySold", "yearSold"), sep = "/") %>%
  mutate(yearSold = strtoi(yearSold) + 2000) %>%
  mutate(daySold = strtoi(daySold), monthSold = strtoi(monthSold)) %>%
  mutate(age = yearSold - strtoi(age)) %>%
  select(- Property.Marketing.Period) %>%
  mutate(originalPrice = strtoi(str_replace_all(str_sub(originalPrice, 2, -4), ",", ""))) %>%
  mutate(lastPrice = strtoi(str_replace_all(str_sub(lastPrice, 2, -4), ",", ""))) %>%
  select(- X) %>%
  mutate(lotSize = as.double(lotSize)) %>%
  select(- Address) %>%
  mutate(marketDays = strtoi(marketDays))

#some final tidying

badStyles <- house %>% 
  group_by(style) %>%
  summarise(count = n()) %>%
  arrange(count) %>%
  head(43)

house <- house %>%
  select(- lotSize) %>%     #too many NAs
  select(- originalPrice) %>%
  select(- lastPrice) %>%  #omit these two predictors because they are obviously way too similar to the response variable
  mutate(nbhd = replace(nbhd, nbhd == "Battelfield Area", "Battlefield Area"), 
         nbhd = replace(nbhd, nbhd == "princeton Ridge", "Princeton Ridge"), 
         nbhd = replace(nbhd, nbhd == "LIttlebrook", "Littlebrook"), 
         nbhd = replace(nbhd, nbhd == "griggs Farm", "Griggs Farm"), 
         nbhd = replace(nbhd, nbhd == "PrettyBrook Area", "Pretty Brook Area"), 
         nbhd = replace(nbhd, nbhd == "Palmer Sq", "Palmer Square"), 
         nbhd = replace(nbhd, nbhd == "Institute", "Institute Area"), 
         nbhd = replace(nbhd, nbhd == "Rriverside", "Riverside"), 
         nbhd = replace(nbhd, nbhd == "Rriverside", "Riverside"), 
         nbhd = replace(nbhd, nbhd == "Carnegie Lake", "Carnegie Lake Area"), 
         nbhd = replace(nbhd, nbhd == "Town", "Township")) %>%     #fix area names
  filter(! nbhd %in% c("Edgerstoune", "Northridge", "Rushbrook", "Russell Estates", 
                       "The Preserve", "Governors Lane", "Princeton Borough", 
                       "Preserve", "Queenston")) %>%  #delete areas with few houses
  filter(! style %in% badStyles$style) %>%  #delete styles with at most 3 houses
  mutate(style = replace(style, style == "Bi-level", "Bi-Level"), 
         style = replace(style, style == "End Unit", "End-Unit"), 
         style = replace(style, style == "Townhouse", "Townhome"))  #fix style names

knitr::kable(list(house %>% 
  group_by(style) %>%
  summarise(meanSoldPrice = mean(soldPrice)) %>%
  arrange(meanSoldPrice) %>%
  head(10),
  house %>% 
  group_by(nbhd) %>%
  summarise(meanSoldPrice = mean(soldPrice)) %>%
  arrange(meanSoldPrice) %>%
  head(10)), caption = "Top ten styles and neighborhoods with highest meanSoldPrice")
```

# Regression Analysis

Our final models are Lasso and PCR. Lasso has the lowest variance, while PCR has the highest R2.

# Discussion and Limitations

Assumptions possibly not met, curse of dimensionality, did not incorporate census data

# Conclusion

# Additional Work

Introduce the other models we tried